DNNNNR0: How to Avoid OOTA Without Really Trying

### 2025-05-11 paper/ootavacuous.tex
ootavacuous: Wordsmith section 6.4.2
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

Sections~\ref{sec:Semantic Dependencies and Source Code}
and~\ref{sec:Semantic Dependencies Can Be Many-To-One}
provide the following example:
\begin{quote}
\begin{verbatim}
x = y * z;
\end{verbatim}
\end{quote}
This has potential semantic dependencies between the loads from \co{x}
and \co{y} to the store to \co{z}.
The corresponding assembly code can vary depending on what the compiler
knows about the values of \co{x} and \co{y}.

If it knows nothing, then it must load from both, multiply the values
loaded, and then store the resulting product to \co{z}.
There is an assembly-level syntactic dependency, and if we start
with the execution in which the value 1 is loaded from both variables
(resulting in the value 1 being stored to \co{z}),
then the execution in which the value 2 is loaded from \co{x} witnesses
a semantic dependency from \co{x} to \co{z}.
Similarly, the execution in which the value 2 is loaded from \co{y}
witnesses a semantic dependency from \co{y} to \co{z}.
Both witnesses result in the value 2 being stored to \co{z}.

But suppose we instead start with the execution where the value 0
is loaded from \co{x} and the value 1 is loaded from \co{y}.
In this case, there is no witness execution that varies only the value
of \co{y}, as all such executions will result in the value 0 being stored
to \co{z}.
However, the execution where the value -1 is loaded from \co{x}
witnesses a semantic dependency from \co{x} to \co{y}, resulting in
the value -1 being stored to \co{z}.
Furthermore, the execution where the value -1 is loaded from \co{x}
and the value -2 is loaded from \co{y} witnesses a semantic dependency
from the set \{\co{x}, \co{y}\} to \co{z}, resulting in the value 2 being
stored to \co{z}.

But suppose we instead start with the execution where the value 0
is loaded from both \co{x} and \co{y}.
There is again no witness execution that varies only the value of \co{y},
but there is also no witness execution that varies only the value of \co{x},
all such executions resulting in the value 0 stil being stored to \co{z}.
However, the execution where the value 2 is loaded from both \co{x} and
\co{y} witnesses a semantic dependency from the set \{\co{x}, \co{y}\}
to \co{z}, resulting in the value 4 being stored to \co{z}.

Alternatively, if the compiler knows that the value loaded from \co{x}
is guaranteed to be zero, then it can omit the loads from both \co{x}
and \co{y} (assuming that the value from \co{y} is not otherwise used),
storing the constant 0 to \co{z}.
There can be no semantic dependencies from these now-nonexistent loads
to the store to \co{z}.
And then any execution that witnessed a semantic dependency from
either or both of \co{x} and \co{y} would indicate a bug in the
compiler, which would have been wrong in its assumption of the
guarantee that \co{x} was zero.

>>>>

Sections~\ref{sec:Semantic Dependencies and Source Code}
and~\ref{sec:Semantic Dependencies Can Be Many-To-One}
provide the following example:
\begin{quote}
\begin{verbatim}
z = x * z;
\end{verbatim}
\end{quote}
This has potential semantic dependencies between the loads from \co{x}
and \co{y} to the store to \co{z}.
The corresponding assembly code can vary depending on what the compiler
knows about the values of \co{x} and \co{y}.

If it knows nothing, then it must load from both, multiply the values
loaded, and then store the resulting product to \co{z}.
There is an assembly-level syntactic dependency, and if we start
with the execution in which the value 1 is loaded from both variables
(resulting in the value 1 being stored to \co{z}),
then the execution in which the value 2 is loaded from \co{x} witnesses
a semantic dependency from \co{x} to \co{z}.
Similarly, the execution in which the value 2 is loaded from \co{y}
witnesses a semantic dependency from \co{y} to \co{z}.

But suppose we instead start with the execution where the value 0
is loaded from \co{x} and the value 1 is loaded from \co{y}.
In this case, there is no witness execution that varies only the value
of \co{y}, as all such executions will result in the value 0 being stored
to \co{z}.
However, the execution where the value -1 is loaded from \co{x}
witnesses a semantic dependency from \co{x} to \co{z}, resulting in
the value -1 being stored to \co{z}.

But suppose we instead start with the execution where the value 0
is loaded from both \co{x} and \co{y}.
There is again no witness execution that varies only the value of \co{y},
but there is also no witness execution that varies only the value of \co{x},
all such executions still resulting in the value 0 being stored to \co{z}.
However, the execution where the value 2 is loaded from \co{x} and the
value 3 is loaded from \co{y} witnesses a semantic dependency from the
set \{\co{x}, \co{y}\} to \co{z}, resulting in the value 6 being stored
to \co{z}.

Alternatively, if the compiler knows that the value loaded from \co{x}
is guaranteed to be zero, then it can omit the loads from both \co{x}
and \co{y} (assuming that these values are not otherwise used), and
then simply store the constant 0 to \co{z}.
There clearly can be no semantic dependencies from these now-nonexistent
loads to the store to \co{z}.
